---
title: "Comparison between Bakta and Prokka for the automatic annotation of prokaryotic genomes"
author: "Alex Huebner"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output:
    github_document:
        fig_width: 8
        fig_height: 6
        toc: true
bibliography: ./221122_Bakta_vs_Prokka.bib
---

In the past ten years, there has been a huge increase in the number of metagenome-assembled genomes
(MAGs) due to improvement in *de novo* assembly tools, such as MEGAHIT [@Li2015] and metaSPAdes
[@Nurk2017], and the decreasing costs for Illumina short-reads sequencing. Many of the
pipelines available for *de novo* assembly of short-read sequencing data and subsequent binning of
the yielded contigs, e.g. nf-core/mag [@Krakau2022] or ATLAS [@Kieser2020], are using the software
Prokka [@SeemanProkka2014] to rapidly annotate the MAGs. 

Due to its functionality, Prokka has become the de facto standard for this type of bacterial genome
annotation. However, it has been developed further in the last years and recently the author of
Prokka, Torsten Seeman, recommended on [Twitter](https://twitter.com/torstenseemann/status/1565471892840259585?s=20&t=SgsPvhR6e09ByVdXs5PIGg)
to use Bakta [@Schwengers2021] as a successor of Prokka.

In contrast to Prokka, Bakta aims to increase the ability to assign the newly annotated coding
sequences to genes that are available in reference databases and to improve the export of the
annotations, e.g. by using JSON files. The authors of Bakta performed a single genome (*E. coli*)
benchmark between Bakta and Prokka, and could show that Bakta had a higher number of proteins with a
known function at a similar run time.
 
To validate these results on a larger set of genomes and to evaluate the suitability of the Bakta's
reference databases, I ran Bakta on the 29 MAGs obtained from EMN001 that have been previously
annotated using Prokka and compared the annotation results with each other.


# Results

```{r libraries, echo=F}
library(knitr)
library(rmarkdown)
library(data.table)
library(tidyverse)
library(pander)
library(patchwork)
opts_chunk$set(echo=F, warning=F, message=F, dpi=150)
theme_set(theme_classic(base_size = 10))
```

```{r load_data}
```

# References
